### 2014-03-01 paper/formal.tex
Improvements to formal.

%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
In our case, |specLang alpha (targetLang nop) = (ifc C, ifc
V, .->, erasef il)|, such that $|ifc C| = |iS| \times \prod |it|$ and
$|ifc V| = |iS| \times |it|_V$, where $|it|_V \subset |it|$ is the
type for tasks whose expressions have been reduced to
values.\footnote{
  Here, we abuse notation by describing types for configuration parts using the
  same metavariables as the ``instance'' of the type, e.g., |it| for the type of
  task.
}
The erased configuration |tyerase (ifc C)| extends |ifc C| with configurations
containing |bullet|; Figure~\ref{fig:erasure} gives the precise definition for
our erasure function |erasef il|.
%
A task and its corresponding message queue is completely removed from the task
list if its label does not flow to the attacker observation level |il|;
otherwise, we apply the erasure function homomorphically and remove any message
from the task's message queue that are more sensitive than |il|.
%}

@@

We note that the definition of |erasef il| precisely captures the IFC
language's attacker model. 
%
For instance, in our case, by removing sensitive tasks and messages from the
system, we assume that that the attacker cannot observe the number of sensitive
tasks or messages.
%
While such assumptions are standard~\tocite{}, our definitions allow for
stronger attackers that may be able to inspect resource usage.\footnote{
  We believe that we can extend |specLang alpha (targetLang nop)| to
  such models using the resource limits techniques of~\tocite{}.
  %
  We leave this extension to future work.
}

Non-interference relies on the definition of an attacker's observational power at
security level |l|, which is typically defined as an equivalence
relation---called |l|-equivalence---on configurations, i.e., an attacker cannot
distinguish two configurations that are |l|-equivalent.
%
Since our erasure function captures the attacker model, we simply define this
equivalence as the syntactic-equivalence of erased configurations.
%
\begin{definition}[|l|-equivalence]
    In a language |(C, .->, erasef l)|, two machine configurations
    |memberf (c, c') C| are considered $l$-equivalent, written as |c ~= c'|,
    if |erase l c = erase il c'|.
\end{definition}
%

Intuitively, we can now state that a language satisfies non-interference if an
attacker at level |l| cannot distinguish the runs of any two |l|-equivalent
configurations.
%
This precise non-interference property is called termination sensitive non-interference
(TSNI), since it considers runs of programs, and thus does not allow for leaks
due termination or internal timing behavior.
%
Formally, we define TSNI as follows.

@@

\begin{theorem}[Concurrent specification language is TSNI]
  \label{thm:rr-tsni}
For any target language |targetLang nop|, |specLang roundrobinf
(targetLang nop)| satisfies TSNI.
\end{theorem}

In general TSNI will not hold for an arbitrary scheduler |alpha|.
%
For example, |specLang alpha targetLang| with a scheduler that inspects a
sensitive threads current state when deciding which thread to schedule next
will clearly not satisfy TSNI.
%
In Section~\toref{} we detail the preceise scheduler for which our
specification language is TSNI.
%
Here, we simply remark that even non-adversairal schedulers are not always safe.

Consider, for example, the sequential scheduling policy |seqf| given in
Figure~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI.
%
Concretely, consider a target language similar to |targetLangML| with an
additional expression terminal |diverge| that denotes a divergent computation,
i.e., |diverge| always reduces to |diverge| and a simple label lattice |{pub,
sec}| such that |pub canFlowTo sec|, but |sec cantFlowTo pub|.  We can write the
following program in the specification language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(when false diverge)),  fullconf 2 pub tS1 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(when true diverge)),   fullconf 2 pub tS2 ie
\end{code}
These two configurations are |pub|-equivalent, however |ic1| will reduce (in two
steps) to |ic1 = iS ; fullconf 2 pub tS1 (IT diverge)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Setion~\ref{}.
} then the |sec| tasks's decision to diverge or not is directly leaked to a
public entitiy.

In certain cases, however, a TSNI-safe concurrent scheduling policy such as
|roundrobinf| cannot be considered without imposig a non-trivial burden on the
concrete implementation.
%
(Section~\ref{ref:real} describes a particular embedding with JavsScript as the
target language.)
%
Hence, we consider an alternative non-interference property---called termination insensitive
non-interference---that can also be upheld by sequential languages at the cost
of leaking one bit through (non)-termination.
%
\begin{definition}[Termination insensitive non-interference (TINI)]
  A language |(C, V, .->, erasef l)| is termination
  insensitive non-interfering if for any label |l|, and configurations
  |c1, c2 member C| and |c1', c2' member V| if
   $|c1| \approx_{|l|} |c2|$,
   $|c1| |.->|^* |c1'|$, and
   $|c2| |.->|^* |c2'|$, then
   $|c1'| \approx_{|l|} |c2'|$.
\end{definition}

TSNI states that if we take two |l|-equivalent configurations, and both
configurations reduce to terminal configuration states (for which there are no
possible further transitions), then the end configurations are also
|l|-equivalent.
%
We highlight that this statement is much weaker than TSNI: it only states that
terminating programs do not leak sensitive data.

Nevertheless, we believe TINI to be a useful for many scenariors and thus we
show that our IFC language with the sequential scheduling policy |seqf|
satisfies TINI.
%
\begin{theorem}[Sequential specification language is TINI]
  \label{thm:seq-tini}
For any target language |targetLang nop|, |specLang seqf (targetLang
nop)| satisfies TINI.
\end{theorem}

>>>>

%{
%format .->  = "\overset{\alpha}{\hookrightarrow}"
Our language |specLang alpha (targetLang nop)| can be
represented as |(ifc C, ifc V, .->, erasef il)|, such that
$|ifc C| = |iS| \times \prod |it|$ and
$|ifc V| = |iS| \times |it|_V$, where $|it|_V \subset |it|$ is the
type for tasks whose expressions have been reduced to
values.\footnote{
  Here, we abuse notation by describing types for configuration parts using the
  same metavariables as the ``instance'' of the type, e.g., |it| for the type of
  task.
}
The erased configuration |tyerase (ifc C)| extends |ifc C| with configurations
containing |bullet|, and Figure~\ref{fig:erasure} gives the precise definition for
our erasure function |erasef il|.
%
A task and its corresponding message queue is completely removed from the task
list if its label does not flow to the attacker observation level |il|.
Otherwise, we apply the erasure function homomorphically and remove any message
from the task's message queue that are more sensitive than |il|.
%}

@@

We note that the definition of |erasef il| captures the IFC
language's attacker model. 
%
In our case, we assume that the attacker cannot observe sensitive tasks or
messages, or even just the number of such entities.
%
While such assumptions are standard~\tocite{}, our definitions allow for
stronger attackers that may be able to inspect resource usage.\footnote{
  We believe that we can extend |specLang alpha (targetLang nop)| to
  such models using the resource limits techniques of~\tocite{}.
  %
  We leave this extension to future work.
}

Non-interference relies on the definition of an attacker's observational power at
security level |l|, which is typically defined as an equivalence
relation---called |l|-equivalence---on configurations.  An attacker
at level |l| cannot
distinguish two configurations that are |l|-equivalent.
%
Since our erasure function captures the attacker model, we simply define this
equivalence as the syntactic-equivalence of erased configurations.
%
\begin{definition}[|l|-equivalence]
    In a language |(C, .->, erasef l)|, two machine configurations
    |memberf (c, c') C| are considered $l$-equivalent, written as |c ~= c'|,
    if |erase l c = erase il c'|.
\end{definition}
%

Intuitively, we can now state that a language satisfies non-interference if an
attacker at level |l| cannot distinguish the runs of any two |l|-equivalent
configurations.
%
This precise non-interference property is called termination sensitive non-interference
(TSNI).  Besides the obvious requirement to not leak secret information
to public channels, this definition also requires the termination
of public tasks to be independent of secret tasks.
%
Formally, we define TSNI as follows.

@@

\begin{theorem}[Concurrent IFC language is TSNI]
  \label{thm:rr-tsni}
For any target language |targetLang|, |specLang roundrobinf
targetLang| satisfies TSNI.
\end{theorem}

In general TSNI will not hold for an arbitrary scheduler |alpha|.
%
For example, |specLang alpha targetLang| with a scheduler that inspects a
sensitive threads current state when deciding which thread to schedule next
will in general not satisfy TSNI.
%
%In Section~\toref{} we detail the preceise scheduler for which our
%specification language is TSNI.
%

However, even non-adversairal schedulers are not always safe.
Consider, for example, the sequential scheduling policy |seqf| given in
Figure~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI.
%
Concretely, consider a target language similar to |targetLangML| with an
additional expression terminal |diverge| that denotes a divergent computation,
i.e., |diverge| always reduces to |diverge| and a simple label lattice |{pub,
sec}| such that |pub canFlowTo sec|, but |sec cantFlowTo pub|.
Consider the following two configurations in this language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(when false diverge)),  fullconf 2 pub tS1 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(when true diverge)),   fullconf 2 pub tS2 ie
\end{code}
These two configurations are |pub|-equivalent, but |ic1| will reduce (in two
steps) to |ic1 = iS ; fullconf 2 pub tS1 (IT diverge)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Setion~\ref{sec:extensions}.}
then the |sec| tasks's decision to diverge or not is directly leaked to a
public entitiy.

% sheule: I feel like this forward reference is not useful here.
%In certain cases, however, a TSNI-safe concurrent scheduling policy such as
%|roundrobinf| cannot be considered without imposing a non-trivial burden on the
%concrete implementation.
%(Section~\ref{ref:real} describes a particular embedding with JavsScript as the
%target language.)
%
To accomodate for sequential languages, or cases where a weaker guarantee
is sufficient, we consider an alternative non-interference property called termination insensitive
non-interference (TINI).  This property can also be upheld by sequential languages at the cost
of leaking one bit through (non)-termination.
%
\begin{definition}[Termination insensitive non-interference (TINI)]
  A language |(C, V, .->, erasef l)| is termination
  insensitive non-interfering if for any label |l|, and configurations
  |c1, c2 member C| and |c1', c2' member V|, it holds that
  \[
    (|c1| \approx_{|l|} |c2|
    \;\;\land\;\;
    |c1| |.->|^* |c1'|
    \;\;\land\;\;
    |c2| |.->|^* |c2'|)
    \implies
    |c1'| \approx_{|l|} |c2'|
  \]
\end{definition}

TSNI states that if we take two |l|-equivalent configurations, and both
configurations reduce to final configurations (i.e.,
configurations for which there are no
possible further transitions), then the end configurations are also
|l|-equivalent.
%
We highlight that this statement is much weaker than TSNI: it only states that
terminating programs do not leak sensitive data, but makes no statement
about non-terminating programs.

Nevertheless, we believe TINI to be a useful for many scenarios and thus we
show that our IFC language with the sequential scheduling policy |seqf|
satisfies TINI.
%
\begin{theorem}[Sequential IFC language is TINI]
  \label{thm:seq-tini}
For any target language |targetLang|, |specLang seqf targetLang| satisfies TINI.
\end{theorem}
